```{r}  
library(Seurat)
library(tidyverse)
install.packages('devtools')
devtools::install_github('immunogenomics/presto')
library(presto) #Efficient Wilcoxon test with FindMarkers
```

```{r}
#Load all neurons object. Data scaling and initial PCA run on hpc
Cervthorlumb_neurons <- readRDS("/Users/roomerb/cervthorlumb_neurons_post.RDS")
```

#Run PCA & clustering
```{r}
#Principal component analysis reporting top 90 PCs
DefaultAssay(Cervthorlumb_neurons) <- "integrated"
Cervthorlumb_neurons <- RunPCA(Cervthorlumb_neurons, npcs = 90);
DimHeatmap(Cervthorlumb_neurons, dims = 1:15, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 31:45, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 46:60, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 61:75, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 76:90, cells = 500, balanced = TRUE)
ElbowPlot(Cervthorlumb_neurons, ndims = 90)

#Find clusters and calculate UMAP for 90 PCs
DefaultAssay(Cervthorlumb_neurons) <- "integrated"
Cervthorlumb_neurons <- FindNeighbors(Cervthorlumb_neurons, reduction = "pca", dims = 1:90)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 0.5)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 1)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 2)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 4)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 6)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 8)
Cervthorlumb_neurons <- RunUMAP(Cervthorlumb_neurons, reduction = "pca", dims = 1:90)
```
#Assessing first clustering
```{r}
#Dimplot for all neurons at 90 PCs grouped by different cluster resolutions, original identity (sample), embryo age (timepoint) and segment (cervical or lumbar)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.0.5", label = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.1", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.2", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.4", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
DefaultAssay(Cervthorlumb_neurons) <- "RNA"
FeaturePlot(Cervthorlumb_neurons, features = "Slc17a6", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc32a1", order = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "orig.ident")
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "timepoint")
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "segment", raster = F)
DimPlot(Cervthorlumb_neurons_ev, reduction = "umap", group.by = "integrated_snn_res.6", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.8", label = T) + NoLegend()

DefaultAssay(Cervthorlumb_neurons) <- "RNA"
FeaturePlot(Cervthorlumb_neurons, features = "Slc17a6", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc32a1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc5a7", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Onecut1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Zfhx3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Neurod2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Nfib", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Avil", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Pirt", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Tlx2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sox10", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sox9", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Plp1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Olig1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nCount_RNA", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nFeature_RNA", order = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T, raster = F)
```
#First subset round, removing obvious DRG neurons,  cluster containing oligo markers, and low nCount clusters. Used initial neurons object integrated by replicate (neurons were removed and then reintegrated amongst replicates), 90PCs.
```{r}

Cervthorlumb_neurons <- SetIdent(Cervthorlumb_neurons, value = "integrated_snn_res.6")
Cervthorlumb_neurons
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
Cervthorlumb_neurons <- subset(Cervthorlumb_neurons, idents = c(
0,
#1, - Oligos
2,
#3, - low nCount
4,
#5, - low nCount
6,
7,
#8, - low nCount
9,
10,
11,
12,
13,
14,
15,
16,
#17, DRG
18,
19,
20,
21,
22,
23,
24,
25,
26,
27,
28,
29,
30,
31,
32,
33,
34,
35,
36,
37,
38,
39,
40,
41,
42,
43,
44,
#45, - low nCount
46,
47,
48,
49,
50,
51,
52,
53,
54,
55,
56,
57,
58,
59,
60,
61,
62,
63,
64,
65,
66,
67,
68,
69,
70,
71,
72,
73,
74,
75,
76,
77,
78,
79,
80,
81,
82,
#83, DRG
84,
85,
86,
87,
88,
89,
90,
91,
92,
93,
94,
95,
96,
97,
98,
99,
100,
101,
102,
103,
104,
105,
106,
107,
108,
109,
110,
111,
112,
113,
#114, Oligos
115,
116,
#117, - low nCount
118,
119,
120,
121,
122,
123,
124,
125,
126,
127,
128,
129,
130,
131,
132,
133,
134,
135,
136,
137,
138,
#139, - low nCount
140,
141,
142))
Cervthorlumb_neurons
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
```
#Rerun PCA & clustering after subset round one
```{r}
#Principal component analysis reporting top 90 PCs
DefaultAssay(Cervthorlumb_neurons) <- "integrated"
Cervthorlumb_neurons <- RunPCA(Cervthorlumb_neurons, npcs = 90);
DimHeatmap(Cervthorlumb_neurons, dims = 1:15, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 31:45, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 46:60, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 61:75, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 76:90, cells = 500, balanced = TRUE)
ElbowPlot(Cervthorlumb_neurons, ndims = 90)

#Find clusters and calculate UMAP for 90 PCs
DefaultAssay(Cervthorlumb_neurons) <- "integrated"
Cervthorlumb_neurons <- FindNeighbors(Cervthorlumb_neurons, reduction = "pca", dims = 1:90)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 0.5)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 1)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 2)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 4)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 6)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 8)
Cervthorlumb_neurons <- RunUMAP(Cervthorlumb_neurons, reduction = "pca", dims = 1:90)
```
#Assessing second clustering
```{r}
#Dimplot for all neurons at 90 PCs grouped by different cluster resolutions, original identity (sample), embryo age (timepoint) and segment (cervical or lumbar)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.0.5", label = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.1", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.2", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.4", label = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
DefaultAssay(Cervthorlumb_neurons) <- "RNA"
FeaturePlot(Cervthorlumb_neurons, features = "Slc17a6", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc32a1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc5a7", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Zfhx3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Nfib", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nCount_RNA", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nFeature_RNA", order = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "orig.ident")
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "timepoint")
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "segment")
DimPlot(Cervthorlumb_neurons_ev, reduction = "umap", group.by = "integrated_snn_res.6", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.8", label = T) + NoLegend()
DefaultAssay(Cervthorlumb_neurons) <- "RNA"
FeaturePlot(Cervthorlumb_neurons, features = "Slc17a6", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc32a1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc5a7", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Onecut1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Zfhx3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Neurod2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Nfib", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Avil", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Pirt", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Tlx2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sox10", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sox9", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Plp1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Olig1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nCount_RNA", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nFeature_RNA", order = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
```
#Second subset round, removing low nCount and possible doublets
```{r}

Cervthorlumb_neurons <- SetIdent(Cervthorlumb_neurons, value = "integrated_snn_res.6")
Cervthorlumb_neurons
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
Cervthorlumb_neurons <- subset(Cervthorlumb_neurons, idents = c(
0,
#1, low nCount
2, 
3,
4,
5,
6,
7,
#8, low nCount
9,
10,
11,
12,
13,
14,
15,
16,
17,
18,
19,
20,
21,
22,
23,
24,
25,
26,
27,
28,
29,
30,
31,
32,
33,
34,
35,
36,
37,
38,
39,
40,
41,
42,
43,
44,
45,
46,
47,
48,
49,
50,
51,
52,
53,
54,
55,
56,
57,
58,
59,
60,
61,
62,
63,
64,
65,
66,
67,
68,
69,
70,
71,
72,
73,
74,
75,
76,
77,
78,
79,
80,
81,
82,
83,
84,
85,
86,
87,
88,
89,
90,
91,
92,
93,
94,
95,
96,
97,
98,
99,
100,
101,
102,
103,
104,
105,
106,
107,
108,
109,
110,
111,
112,
113,
114,
115,
116,
117,
118,
119,
120,
121,
122,
123,
#124, low nCount
#125, possible doublets
126,
127,
128,
129,
130,
131,
132,
133,
134,
135,
136,
137))
Cervthorlumb_neurons
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
```
#Rerun PCA & clustering after subset round two
```{r}
#Principal component analysis reporting top 90 PCs
DefaultAssay(Cervthorlumb_neurons) <- "integrated"
Cervthorlumb_neurons <- RunPCA(Cervthorlumb_neurons, npcs = 90);
DimHeatmap(Cervthorlumb_neurons, dims = 1:15, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 31:45, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 46:60, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 61:75, cells = 500, balanced = TRUE)
DimHeatmap(Cervthorlumb_neurons, dims = 76:90, cells = 500, balanced = TRUE)
ElbowPlot(Cervthorlumb_neurons, ndims = 90)
#Find clusters and calculate UMAP for 90 PCs
DefaultAssay(Cervthorlumb_neurons) <- "integrated"
Cervthorlumb_neurons <- FindNeighbors(Cervthorlumb_neurons, reduction = "pca", dims = 1:90)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 0.5)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 1)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 2)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 4)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 6)
Cervthorlumb_neurons <- FindClusters(Cervthorlumb_neurons, resolution = 8)
Cervthorlumb_neurons <- RunUMAP(Cervthorlumb_neurons, reduction = "pca", dims = 1:90)
```
#Assessing third clustering. Third version of spinal neuron object; used hereafter
```{r}
#Dimplot for all neurons at 90 PCs grouped by different cluster resolutions, original identity (sample), embryo age (timepoint) and segment (cervical or lumbar)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.0.5", label = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.1", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.2", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.4", label = T) + NoLegend()
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.6", label = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "integrated_snn_res.8", label = T)
DefaultAssay(Cervthorlumb_neurons) <- "RNA"
FeaturePlot(Cervthorlumb_neurons, features = "Slc17a6", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc32a1", order = T)
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "orig.ident")
DimPlot(Cervthorlumb_neurons, reduction = "umap", group.by = "segment")

DefaultAssay(Cervthorlumb_neurons) <- "RNA"
FeaturePlot(Cervthorlumb_neurons, features = "Fshr", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Defa29", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Gm27246", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Gm5083", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Krt71", order = T)

FeaturePlot(Cervthorlumb_neurons, features = "Slc17a6", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc32a1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Slc5a7", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sox2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Onecut1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Zfhx3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Neurod2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Nfib", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nCount_RNA", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "nFeature_RNA", order = T)

FeaturePlot(Cervthorlumb_neurons, features = "Onecut1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Zfhx3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Neurod2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Nfib", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Barhl1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Foxd3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Isl1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Lmx1b", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Evx1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Vsx2", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sim1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Tfap2b", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Dmrt3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Gata3", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "En1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Pkd2l1", order = T)
FeaturePlot(Cervthorlumb_neurons, features = "Sox2", order = T)

FeaturePlot(Cervthorlumb_neurons, features = "Zic5", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Zic2", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Zic3", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Zic4", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Zic1", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Pax2", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Tfap2b", order = T, raster = F)
FeaturePlot(Cervthorlumb_neurons, features = "Satb2", order = T, raster = F)
```

```{r}
#For now, we will keep this size object, and this will be the version loaded in the future.
#Save neurons object - currently saved post subset
saveRDS(Cervthorlumb_neurons, "/Users/roomerb/cervthorlumb_neurons_post.RDS")
```